---
title: "Descriptive statistics by urban/rural area"
author: "Maggie Walters"
date: "June 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work/")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))
```

## ANOVA with RMSP as a factor
```{r rmsp as a factor}
#list of each county
county_vec <- as.character(unique(data$COUNTY))
#create blank upper quartile matrix
upperq_matrix <- matrix(rep(NA, length(county_vec) * 12), nrow = 12, ncol = length(county_vec))
colnames(upperq_matrix) <- county_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(upperq_matrix) <- months

#find upper quartile for each county in each month
for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:387){
    which.county_j <- which(x$COUNTY == county_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    upperq_matrix[i,j] <- quantile(county_j_ages)[4]
  }
}
upper_matrix <- matrix(rep(NA, 4 * 12*length(county_vec)), ncol = 4)
colnames(upper_matrix) <- c("UPPER", "COUNTY", "MONTH", "RMSP")

#fill in upper
for(i in 1:12 * length(county_vec)){
  upper_matrix[,1] <- upperq_matrix[,]
}
#fill in months
upper_matrix[,3] <- rep(seq(1:12), length(county_vec))
#fill in counties
upper_matrix[seq(1,12),2] <- rep(county_vec[1], 12)
for(i in 1:49){
  x <- 12 * (i-1)
  upper_matrix[x + seq(1,12),2] <- rep(county_vec[i], 12)
}
#convert to data frame
upper_matrix <- as.data.frame(upper_matrix)

#isolate RMSPs
x <- rep(NA, length(county_vec))
for(i in 1:length(county_vec)){
  which.i <- which(data$COUNTY == county_vec[i])
  y <- rep(NA, length(which.i))
  y <- data$RMSP[which.i]
  x[i] <- mean(y)
}
rmsp_vec <- x

#fill in RMSPs
upper_matrix[seq(1,12),4] <- rep(rmsp_vec[1], 12)
for(i in 1:length(county_vec)){
  x <- 12 * (i-1)
  upper_matrix[x + seq(1,12),4] <- rep(rmsp_vec[i], 12)
}

#remove NAs
upper_matrix <- subset(upper_matrix, !is.na(UPPER))

#change RMSP to a factor
upper_matrix$RMSP <- as.factor(upper_matrix$RMSP)

#change UPPER to a number
upper_matrix$UPPER <- as.numeric(upper_matrix$UPPER)

##rmsp_aov_mod <- aov(UPPER ~ RMSP, data = upper_matrix)

upper_mean_is <- subset(upper_matrix, upper_matrix$RMSP == 1)
y1 <- mean(upper_mean_is$UPPER)
x1 <- sd(upper_mean_is$UPPER)

upper_mean_not <- subset(upper_matrix, upper_matrix$RMSP == 0)
y2 <- mean(upper_mean_not$UPPER)
x2 <- sd(upper_mean_not$UPPER)

#sort by month for both
x <- order(upper_mean_is$MONTH)
upper_mean_is <- upper_mean_is[order(upper_mean_is$MONTH),]

```

### Preliminary analysis

* rmsp_aov_mod is not displaying a p-value, only sum of squares and degrees of freedom. 
* In order to provide some sort of descriptive statistics here, the mean of those within the regional metropolitan area (RMSP = 1) is `r y1` and the standard deviation is `r x1`.
* The mean of those not within the regional metropolitan area (RMSP = 0) is `r y2` and the standard deviation is `r x2`. 


**TO DO:**

* Sort by month regardless of county, so that progress of the epidemic within each county can be accounted for. 
* Plot upper_mean_is (black) and upper_mean_not (red) in order to look for a visual trend/sign that there is a difference between counties within the regional metropolitan area and not within the regional metropolitan area. 

## Divide counties by RMSP

```{r Split between 0 and 1 for RMSP}
#list of each county
county_vec <- as.character(unique(data$COUNTY))

#are not in regional metropolitan area
rural <- subset(data, data$RMSP == 0)

#are in the regional metropolitan area
urban <- subset(data, data$RMSP == 1)

```

## Find upper quartile for rural

```{r pressure, echo=FALSE}
length.rural <- dim(rural)[1]
unique.rural <- unique(rural$COUNTY)

####################################
#create upper matrix
####################################
#create blank upper quartile matrix
upper.rural.mat <- matrix(rep(NA, length(unique.rural) * 12), nrow = 12, ncol = length(unique.rural))
colnames(upper.rural.mat) <- unique.rural
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(upper.rural.mat) <- months

#find upper quartile for each county in each month
for(i in 1:12){
  x <- subset(rural, rural$MONTH == i)
  for(j in 1:length(unique.rural)){
    which.county_j <- which(x$COUNTY == unique.rural[j])
    county_j_ages <- x$AGE[which.county_j]
    upper.rural.mat[i,j] <- quantile(county_j_ages)[4]
  }
}

####################################
#create blank upper quartile matrix
####################################
#create matrix
upper_rural <- matrix(rep(NA, length(unique.rural) * 3), nrow = length(unique.rural) * 12, ncol = 3)
colnames(upper_rural) <- c("UPPER", "COUNTY", "MONTH")
#fill in month
upper_rural[,3] <- rep(seq(1:12), length(unique.rural))
#fill in upper quartile
for(i in 1:12 * 348){
  upper_rural[,1] <- upper.rural.mat[,]
}
#fill in counties
unique.rural <- as.character(unique.rural)
upper_rural[1:12,2] <- rep(unique.rural[1], 12)
for(i in 1:length(unique.rural)){
  x <- 12 * (i-1)
  upper_rural[x + seq(1,12),2] <- rep(unique.rural[i], 12)
}
#remove all NAs
upper_rural <- as.data.frame(upper_rural)
upper_rural <- subset(upper_rural, !is.na(upper_rural$UPPER))
```

## Run ANOVA for county/month: UPPER 

```{r ANOVA for upper and county/month}
upper_rural$UPPER <- as.numeric(upper_rural$UPPER)

####################################
#COUNTY
####################################
#p = 0.0322
upper_county_aov <- aov(UPPER ~ COUNTY, data = upper_rural)
#p = 6.897e-10, not normal
shapiro.test((resid(upper_county_aov)))
plot(upper_county_aov, which = c(1,2))
hist(resid(upper_county_aov))

#########################
#COUNTY: NON-PARAMETRIC
#########################
#KW-test: p = 0.0747
upper_county_kw <- kruskal.test(UPPER ~ COUNTY, data = upper_rural)




####################################
#MONTH
####################################
#p = 0.0555
upper_month_aov <- aov(UPPER ~ MONTH, data = upper_rural)
#p = 2.242e-14, not normal
shapiro.test((resid(upper_month_aov)))
plot(upper_month_aov, which = c(1,2))
hist(resid(upper_month_aov))

#########################
#MONTH: NON-PARAMETRIC
#########################
#KW-test: p = 0.07786
upper_month_kw <- kruskal.test(UPPER ~ MONTH, data = upper_rural)




####################################
#TWO-WAY
####################################
#p = 0.0322
upper_twoway_aov <- aov(UPPER ~ COUNTY * MONTH, data = upper_rural)
#p = 6.897e-10, not normal
shapiro.test((resid(upper_county_aov)))
plot(upper_county_aov, which = c(1,2))
hist(resid(upper_county_aov))

```
### Preliminary results

* Tried many transformations, but could not normalize data. Attempted a KW test for month and county. 
* Just realized that I should do a ANOVA with upper quartile and then the factor being RMSP. Need to use all data. 
