---
title: "age_distribution_of_cases"
author: "Maggie Walters"
date: "June 22, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work/")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))

county_vec <- as.character(unique(data$COUNTY))
```

## Document Synopsis

This document aims to visually and quantitatively describe the different age distribution of cases in the counties represented in spm.data.long.csv. Densities of cases for age groups will be calculated and then compared between counties. 

The difference between counties will be used in order to explore possible predictive relationships between population size or urbanness of a county and it's density of cases in different age classes. 

## Age distribution scatter plots by county

Scatter plots will be created in order to visually identify the age groups which frequently have a low density of cases. Depending on results, this may be limited to counties with number of cases above a certain threshold. 

```{r age distribution scatter plots by county}
for(i in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[i])
  if(length(subset_county$AGE) > 20){
    plot(subset_county$WEEK, subset_county$AGE, pch = 20,
         main = county_vec[i], xlab = "WEEK", ylab = "AGE")
  }
}

```

### Preliminary Analysis

## Confidence intervals for age group densities

Created two_matrix, three_matrix, four_matrix, and five_matrix which contain the density of cases in age groups of the corresponding span for each county. Density was just calculated as the number of cases in that specific class divided by the number of cases in that county total.  

```{r estimate confidence intervals for age group densities}
#find appropriate age windows, look at age groups with ranges ranging from 2 to 5 years
density_age <- function(max.age, size){
  amount <- round(max.age / size)
  amount.vec <- rep(NA, amount)
  amount.vec <- as.data.frame(amount.vec)
  amount.vec[1,] <- length(which(subset_county$AGE <= size))
  for(i in 2:amount){
    before <- as.integer(sum(amount.vec[1: i - 1,]))
    amount.vec[i,] <- length(which(subset_county$AGE <= (size * i))) - before
  }
  density.vec <- amount.vec / length(subset_county$AGE)
  return(density.vec)
}

#matrix for age group of 2's
two_matrix <- matrix(rep(NA, round(84/2) * length(county_vec)), ncol = length(county_vec))
two_matrix <- as.data.frame(two_matrix)
colnames(two_matrix) <- county_vec
row.names(two_matrix) <- c("<2", "2-4", "4-6", "6-8", "8-10",
                           "10-12", "12-14", "14-16", "16-18", "18-20",
                           "20-22", "22-24", "24-26", "26-28", "28-30",
                           "30-32", "32-34", "34-36", "36-38", "38-40",
                           "40-42", "42-44", "44-46", "46-48", "48-50",
                           "50-52", "52-54", "54-56", "56-58", "58-60",
                           "60-62", "62-64", "64-66", "66-68", "68-70",
                           "70-72", "72-74", "74-76", "76-78", "78-80",
                           "80-82", "82-84")

for(j in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[j])
  max.age <- max(subset_county$AGE)
  density <- unlist(density_age(max.age, 2))
  length_d <- length(density)
  density <-  c(density, rep(NA, (42 - length_d)))
  two_matrix[,j] <- density
}
#matrix for age groups of 3's
three_matrix <- matrix(rep(NA, round(84/3) * length(county_vec)), ncol = length(county_vec))
three_matrix <- as.data.frame(three_matrix)
colnames(three_matrix) <- county_vec
rownames(three_matrix) <- c("<3", "3-6", "6-9",
                             "9-12", "12-15", "15-18",
                             "18-21", "21-24", "24-27",
                             "27-30", "30-33", "33-36",
                             "36-39", "39-42", "42-45",
                             "45-48", "48-51", "51-54",
                             "54-57", "57-60", "60-63",
                             "63-66", "66-69", "69-72",
                             "72-75", "75-78", "78-81",
                             "81-84")

for(j in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[j])
  max.age <- max(subset_county$AGE)
  density <- unlist(density_age(max.age, 3))
  length_d <- length(density)
  density <-  c(density, rep(NA, (28 - length_d)))
  three_matrix[,j] <- density
}
#matrix for age groups of 4's
four_matrix <- matrix(rep(NA, 21 * length(county_vec)), ncol = length(county_vec))
four_matrix <- as.data.frame(four_matrix)
colnames(four_matrix) <- county_vec
row.names(four_matrix) <- c("<4", "4-8", "8-12", "12-16",
                            "16-20", "20-24", "24-28", "28-32",
                            "32-36", "36-40", "40-44", "44-48",
                            "48-52", "52-56", "56-60", "60-64",
                            "64-68", "68-72", "72-76", "76-80",
                            "80-84")

for(j in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[j])
  max.age <- max(subset_county$AGE)
  density <- unlist(density_age(max.age, 4))
  length_d <- length(density)
  density <-  c(density, rep(NA, (21 - length_d)))
  four_matrix[,j] <- density
}
#matrix for age groups of 5's
five_matrix <- matrix(rep(NA, round(84/5) * length(county_vec)), ncol = length(county_vec))
five_matrix <- as.data.frame(five_matrix)
colnames(five_matrix) <- county_vec
row.names(five_matrix) <- c("<5", "5-10", "10-15", "15-20",
                           "20-25", "25-30", "30-35", "35-40",
                           "40-45", "45-50", "50-55", "55-60",
                           "60-65", "65-70", "70-75", "75-80",
                           "80-85")

for(j in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[j])
  max.age <- max(subset_county$AGE)
  density <- unlist(density_age(max.age, 5))
  length_d <- length(density)
  density <-  c(density, rep(NA, (17 - length_d)))
  five_matrix[,j] <- density
}
```

### Preliminary Analysis

It appears that the size of the window does have a siginficant (colloquially) effect on the density in that close age classes have pretty different densities. Going to compare densities on a line graph. 

## Comparison of windows

```{r Comparison of windows}
#x axis for twos
x_two <- rep(NA, 42)
x_two[1] <- 1
for(i in 2:42){
  x_add <- seq(1:41)
  x_two[i] <- i + x_add[i - 1]
}
#x axis for threes
x_three <- rep(NA, 28)
x_three[1] <- 2
for(i in 2:28){
  x_add <- seq(1:28)
  x_three[i] <- 2 * i + x_add[i] -1
}
#x axis for fours
x_four <- rep(NA, 21)
x_four[1] <- 2.5
for(i in 2:21){
  x_add <- seq(1:21)
  x_four[i] <- 3 * i + x_add[i] + 0.5 -1
}
#x axis for fives
x_five <- rep(NA, 17)
x_five[1] <- 3
for(i in 2:17){
  x_add <- seq(1:17)
  x_five[i] <- 4 * i + x_add[i] -2
}

for(i in 1:length(county_vec)){
  subset_county <- subset(data, data$COUNTY == county_vec[i])
  if(length(subset_county$AGE) > 20){
    {plot(x_two, two_matrix[,i], type = "l", ylim = c(0, 0.3), main = county_vec[i])
     lines(x_three, three_matrix[,i], col = "red")
     lines(x_four, four_matrix[,i], col = "blue")
     lines(x_five, five_matrix[,i], col = "green")}
  }
}

```

```{r descriptive stats of density}
#two
mean_two <- rep(NA, 42)
sd_two <- rep(NA, 42)
for(i in 1:42){
  mean_two[i] <- rowMeans(two_matrix[i,], na.rm = TRUE)
  sd_two[i] <- apply(two_matrix, 1, sd, na.rm = TRUE)[i]
}

#three
mean_three <- rep(NA, 28)
sd_three <- rep(NA, 28)
for(i in 1:28){
  mean_three[i] <- rowMeans(three_matrix[i,], na.rm = TRUE)
  sd_three[i] <- apply(three_matrix, 1, sd, na.rm = TRUE)[i]
}

#four
mean_four <- rep(NA, 21)
sd_four <- rep(NA, 21)
for(i in 1:21){
  mean_four[i] <- rowMeans(four_matrix[i,], na.rm = TRUE)
  sd_four[i] <- apply(four_matrix, 1, sd, na.rm = TRUE)[i]
}

#five
mean_five <- rep(NA, 17)
sd_five <- rep(NA, 17)
for(i in 1:17){
  mean_five[i] <- rowMeans(five_matrix[i,], na.rm = TRUE)
  sd_five[i] <- apply(five_matrix, 1, sd, na.rm = TRUE)[i]
}

```

```{r rural and urban}
#RMSP = 1
urban <- subset(data, data$RMSP == "1")
urban_counties <- as.character(unique(urban$COUNTY))

which.urban <- rep(NA, length(urban_counties))
for(i in 1:length(urban_counties)){
 which.urban[i] <- which(county_vec[] == urban_counties[i])
}
two_urban_mat <- two_matrix[which.urban]
three_urban_mat <- three_matrix[which.urban]
four_urban_mat <- three_matrix[which.urban]
five_urban_mat <- five_matrix[which.urban]


#RMSP = 0 
rural <- subset(data, data$RMSP == "0")
rural_counties <- as.character(unique(rural$COUNTY))

which.rural <- rep(NA, length(rural_counties))
for(i in 1:length(rural_counties)){
  which.rural[i] <- which(county_vec[] == rural_counties[i])
}
two_rural_mat <- two_matrix[which.rural]
two_rural <- matrix(rep(NA, 3 * 42 * length(rural_counties)), ncol = 3)
colnames(two_rural) <- c("COUNTY", "CLASS", "DENSITY")
two_rural <- as.data.frame(two_rural)
#fill counties
two_rural[,1] <- rep(rural_counties, 42)

#fill age class
two_rural[seq(1:length(rural_counties)),2] <- x_two[1]
for(i in 1:length(x_two)){
  x <- length(rural_counties) * i + seq(1:length(rural_counties))
  two_rural[x,2] <- x_two[i+1]
}

#fill density
for(i in 1:length(rural_counties)){
  two_rural[i,3] <- two_rural_mat[1,i]
}
for(i in 1:length(rural_counties)){
  for(j in 2:length(x_two)){
    two_rural[(i + 348 * (j - 1) ),3]  <- two_rural_mat[j,i]
  }
  }
two_rural$CLASS <- as.factor(two_rural$CLASS)
two_rural$COUNTY <- as.factor(two_rural$COUNTY)


three_rural_mat <- three_matrix[which.rural]
four_rural_mat <- three_matrix[which.rural]
five_rural_mat <- five_matrix[which.rural]

```
### Preliminary analysis
Created a method (via for loops) to make a matrix which is compatible with ANOVAs. 

Up next: 

* Look for correlation or association between denstiy and class on both a rural and urban level. 
    + This may be best to do with the five_rural_mat and five_urban_mat because they have fewer factor levels. Starting with urban may also be more promising because it will have fewer factors for county. 
* Look into whether the moving windows/data smoothing can be used for the windows of age groups. Try to understand how this works. 
* Plot row means with row sd error bars for the different windows. 



```{r ANOVAs}

```

