---
title: "Interquartile Range of Age of Infection"
author: "Maggie Walters"
date: "June 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
#markdown set up
knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))
```

## Interquartile Range of Age of Infection within Counties

Box-Whisker diagrams were created for each county to graphically examine the interquartile range of age of infection for each county. This will allow for eventual comparison of interquartile ranges among counties. 


```{r Interquartile range of aoi within counties, echo=TRUE}
#list of each county
county_vec <- as.character(unique(data$COUNTY))

#create blank IQR matrix
IQR_matrix <- matrix(rep(NA, length(county_vec)), nrow = 1, ncol = length(county_vec))
colnames(IQR_matrix) <- county_vec

#find IQR for each county 
for(i in 1:length(county_vec)){
  which.county_i <- which(data$COUNTY == county_vec[i])
  age_i <- data$AGE[which.county_i]
  IQR_matrix[i] <- IQR(age_i)
}

#I don't like the way that this displays but it won't allow me to use View()
#print(IQR_matrix)

#box-whisker plots for each county
for(i in 1:length(county_vec)){
  data_i <- subset(data, data$COUNTY == county_vec[i])
  #par(mfrow = c(3, 3), pty = "s"), want to do a 3 x 3 with nine plots on one page
  boxplot(data_i$AGE, data = data_i) 
  #add a legend argument which displays sample size (n) for each county: 
  # data$COUNTY == length(county_vec[i])
}

```

### Preliminary Findings

* Maximum interquartile range: `r county_vec[217]`--- `r IQR_matrix[217]`
* Minimum interquartile range: `r county_vec[40]`--- `r IQR_matrix[40]`

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
