---
title: "Interquartile Range of Age of Infection"
author: "Maggie Walters"
date: "June 14, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
#markdown set up
knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work/")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))
```

## Interquartile Range of Age of Infection within Counties

Box-whisker diagrams were created for each county to graphically examine the interquartile range of age of infection for each county. This will allow for eventual comparison of interquartile ranges among counties. 

```{r Interquartile range of aoi within counties, echo=TRUE}
#list of each county
county_vec <- as.character(unique(data$COUNTY))

#create blank lower quartile matrix
lowerq_matrix <- matrix(rep(NA, length(county_vec) * 12), nrow = 12, ncol = length(county_vec))
colnames(lowerq_matrix) <- county_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(lowerq_matrix) <- months

#find lower quartile for each county in each month
for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:387){
    which.county_j <- which(x$COUNTY == county_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    lowerq_matrix[i,j] <- quantile(county_j_ages)[2]
  }
}

#create blank upper quartile matrix
upperq_matrix <- matrix(rep(NA, length(county_vec) * 12), nrow = 12, ncol = length(county_vec))
colnames(upperq_matrix) <- county_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(upperq_matrix) <- months

#find upper quartile for each county in each month
for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:387){
    which.county_j <- which(x$COUNTY == county_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    upperq_matrix[i,j] <- quantile(county_j_ages)[4]
  }
}

#IQR matrix
IQR_matrix <- matrix(rep(NA, length(county_vec) * 12), nrow = 12, ncol = length(county_vec))
colnames(IQR_matrix) <- county_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(IQR_matrix) <- months

#find IQR for each county in each month
for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:387){
    which.county_j <- which(x$COUNTY == county_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    IQR_matrix[i,j] <- IQR(county_j_ages)
  }
}

#box and whisker plot for whole dataset by month
boxplot(data$AGE~data$MONTH, data = data, xlab="Month", 
        ylab="Age", main="Sao Paulo State", las=1)


```

### Preliminary Findings

* The minimum interquartile range was zero, but this would just correspond to one case within a month in a county.
* The maximum interquartile range was `r max(IQR_matrix, na.rm = TRUE)`.
* The average lower quartile was `r mean(lowerq_matrix, na.rm = TRUE)`.
    + Standard deviation was `r sd(lowerq_matrix, na.rm = TRUE)`. 
* The average upper quartile was `r mean(upperq_matrix, na.rm = TRUE)`.
    + Standard deviation was `r sd(upperq_matrix, na.rm = TRUE)`. 
    
Considering doing an ANOVA between counties for both the upper and lower quantiles per month between counties in order to see between which counties they significantly differ. 

**Having trouble creating a useful/ efficient boxplot visualization.** There are 387 counties, so I believe that I need to parse down the number of counties that I am making boxplots for. In essence, it makes no sense to make boxplots for counties with only one case. It also doesn't appear to be useful to create boxplots for counties with 5 cases, because the minimum, lower quartile, mean, upper quartile, and maximum would inherently account for all 5 cases, rather than truly representing the data. 

Re: the box-whisker plots--- the markdown file that Annakate gave me has them for each week of the epidemic (*rather than for each month*) for

* The whole of the Sao Paulo state (i.e. all of the cases in the dataset)
* All of the cases that occurred in the Sao Paulo county (~22,000 cases)

## Number of cases within counties 

While one could probably be assured that all of the cases within the Sao Paulo region are related as the number of measles cases during this period far exceeded what was expected, **is there a minimum number of cases at a county-level needed to be considered an outbreak?**

```{r Number of cases needed to be an outbreak}
#lengths of county-level data
length_mat <- matrix(rep(NA, length(county_vec)), nrow = 1)
colnames(length_mat) <- county_vec

for(i in 1:length(1:387)){
  length_mat[i] <- length(which(data$COUNTY == county_vec[i]))
}
```


### Preliminary Findings

* Number of counties with more than one case: `r length(which(length_mat != 1))`
* Number of counties with more than five cases:  `r length(which(length_mat > 5))`
* Number of counties with more than twenty cases: `r length(which(length_mat > 20))`

```{r Counties with greater than 50 cases, echo=TRUE, fig.width=5, fig.height=5,fig.show="hold"}
#create dataframe for counties with >50 cases
#as.data.frame(length_mat)
which.big_counties <- which(length_mat > 50)
big_counties_vec <- county_vec[which.big_counties]


for(i in 1:length(big_counties_vec)){
  x <- subset(data, data$COUNTY == big_counties_vec[i])
  boxplot(x$AGE ~ x$MONTH, data = x, main = big_counties_vec[i])
  legend("topright", legend = length_mat[which.big_counties[i]])
}


```

## ANOVA using county and month as factors
```{r IQR examination for counties >50}
IQR_matrix_new <- matrix(rep(NA, length(big_counties_vec) * 12), 
                         nrow = 12, ncol = length(big_counties_vec))
colnames(IQR_matrix_new) <- big_counties_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(IQR_matrix_new) <- months

#find IQR for each county >20 in each month
for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:length(big_counties_vec)){
    y <- subset(x, x$COUNTY == big_counties_vec[j])
    IQR_matrix_new[i,j] <- IQR(y$AGE)
  }
}

#ANOVA matrix
anova_mat <- matrix(data = NA, nrow = 49 * 12, ncol = 3)
colnames(anova_mat) <- c("IQR", "COUNTY", "MONTH")
#fill IQR
for(i in 1:49 * 12){
  anova_mat[,1] <- IQR_matrix_new[,]
}
#fill months
anova_mat[,3] <- rep(seq(1,12), length(big_counties_vec))
#fill counties
anova_mat[seq(1,12),2] <- rep(big_counties_vec[1], 12)
for(i in 1:49){
  x <- 12 * (i-1)
  anova_mat[x + seq(1,12),2] <- rep(big_counties_vec[i], 12)
}

anova_mat <- as.data.frame(anova_mat)
anova_mat <- subset(anova_mat, !is.na(anova_mat$IQR))
anova_mat$MONTH <- factor(anova_mat$MONTH)
anova_mat$COUNTY <- as.character(anova_mat$COUNTY)
anova_mat$IQR <- as.numeric(as.character(anova_mat$IQR))

#remove "-" because it's confusing R
which(anova_mat$COUNTY == "EMBU-GUACU")
anova_mat[180,2] <- "EMBU GUACU"
anova_mat[181,2] <- "EMBU GUACU"
anova_mat[182,2] <- "EMBU GUACU"
anova_mat[183,2] <- "EMBU GUACU"
anova_mat$COUNTY <- factor(anova_mat$COUNTY)

#one way between month and IQR, p = 0.0203
anova_IQR_county <- aov(IQR ~ COUNTY, data = anova_mat)
#p = 5.896e-5
shapiro.test((resid(anova_IQR_county)))
plot(anova_IQR_county, which = c(1,2))
hist(resid(anova_IQR_county))

##TRANSFORMATION
IQR_x <- 1 / (anova_mat$IQR + 1)
x_mat <- aov(IQR_x ~ COUNTY, data = anova_mat)
shapiro.test((resid(x_mat)))
hist(resid(x_mat))

#one way between county and IQR, p = <2 e -16
anova_IQR_month <- aov(IQR ~ MONTH, data = anova_mat)
#p = 0.0006897
shapiro.test((resid(anova_IQR_month)))
plot(anova_IQR_month, which = c(1,2))
hist(resid(anova_IQR_month))


#two way with county and IQR, not displaying p value
anova_two_way <- aov(IQR ~ COUNTY + MONTH + COUNTY:MONTH, data = anova_mat)

```
### Preliminary Findings

* There are `r length(big_counties_vec)` counties with more than 20 cases.
* The null hypothesis that county and IQR are unrelated can be rejected with p = 0.00513.
* The null hypothesis that month and IQR are unrelated can be rejected with p = <2e-16.

**TO DO:**

* Either attain normality or complete the non-parametric equivalent of a one-way ANOVA (Kruskal-Wallis)
* Run a two-way ANOVA using MONTH and COUNTY.
    + There is a degrees of freedom issue with the two way anova. There are too many factor levels for the number of observations (634). In order to lower the number of levels I would need to increase the number of cases to be considered from 20 to maybe 30?
    + Would need to do counties with greater than 50 cases (which would amount to 49 cases)
        + Unfortunately, this also reduces the number of observations to 401 (while df is 600)
* Run *a posteri* multiple comparisons test. 

##Differences in 75% percentile 
```{r 75% differences}
upper_matrix_new <- matrix(rep(NA, length(big_counties_vec) * 12), 
                         nrow = 12, ncol = length(big_counties_vec))
colnames(upper_matrix_new) <- big_counties_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(upper_matrix_new) <- months

for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:length(big_counties_vec)){
    which.county_j <- which(x$COUNTY == big_counties_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    upper_matrix_new[i,j] <- quantile(county_j_ages)[4]
  }
}

upper_matrix <- matrix(rep(NA, 3 * 12*length(big_counties_vec)), ncol = 3)
colnames(upper_matrix) <- c("UPPER", "COUNTY", "MONTH")

#fill in upper
for(i in 1:12 * length(big_counties_vec)){
  upper_matrix[,1] <- upper_matrix_new[,]
}

#fill in months
upper_matrix[,3] <- rep(seq(1:12), length(big_counties_vec))

#fill in counties
upper_matrix[seq(1,12),2] <- rep(big_counties_vec[1], 12)
for(i in 1:49){
  x <- 12 * (i-1)
  upper_matrix[x + seq(1,12),2] <- rep(big_counties_vec[i], 12)
}

#convert to data frame
upper_matrix <- as.data.frame(upper_matrix)

#remove NAs
upper_matrix <- subset(upper_matrix, !is.na(UPPER))

#change upper from factor to number
upper_matrix$UPPER <- as.numeric(upper_matrix$UPPER)

#create ANOVA-county model and assess for normality
upper_anova_c <- aov(UPPER ~ COUNTY, data = upper_matrix)
#p = 0.0006119
shapiro.test((resid(upper_anova_c)))
plot(upper_anova_c, which = c(1,2))
hist(resid(upper_anova_c))

#create ANOVA-month model and assess for normality
upper_anova_m <- aov(UPPER ~ MONTH, data = upper_matrix)
#p = 0.0006119
shapiro.test((resid(upper_anova_m)))
plot(upper_anova_m, which = c(1,2))
hist(resid(upper_anova_m))

```
### Preliminary Findings

The upper quartile values were not normal, so it is necessary to either try to find a transformation or consider a non-parametric test (such as the Kruskall-Wallis test).

**Also to be considered are other ways to compare age distributions.**

```{r 25% differences}
lower_matrix_new <- matrix(rep(NA, length(big_counties_vec) * 12), 
                         nrow = 12, ncol = length(big_counties_vec))
colnames(lower_matrix_new) <- big_counties_vec
months <- c("January", "Febuary", "March", 
            "April", "May", "June",
            "July", "August", "September",
            "October", "November", "December")
row.names(lower_matrix_new) <- months

for(i in 1:12){
  x <- subset(data, data$MONTH == i)
  for(j in 1:length(big_counties_vec)){
    which.county_j <- which(x$COUNTY == big_counties_vec[j])
    county_j_ages <- x$AGE[which.county_j]
    lower_matrix_new[i,j] <- quantile(county_j_ages)[2]
  }
}

lower_matrix <- matrix(rep(NA, 3 * 12*length(big_counties_vec)), ncol = 3)
colnames(lower_matrix) <- c("LOWER", "COUNTY", "MONTH")

#fill in lower
for(i in 1:12 * length(big_counties_vec)){
  lower_matrix[,1] <- lower_matrix_new[,]
}

#fill in months
lower_matrix[,3] <- rep(seq(1:12), length(big_counties_vec))

#fill in counties
lower_matrix[seq(1,12),2] <- rep(big_counties_vec[1], 12)
for(i in 1:49){
  x <- 12 * (i-1)
  lower_matrix[x + seq(1,12),2] <- rep(big_counties_vec[i], 12)
}

#convert to data frame
lower_matrix <- as.data.frame(lower_matrix)

#remove NAs
lower_matrix <- subset(lower_matrix, !is.na(LOWER))

#change upper from factor to number
lower_matrix$LOWER <- as.numeric(lower_matrix$LOWER)

#create ANOVA-county model and assess for normality
lower_anova_c <- aov(LOWER ~ COUNTY, data = lower_matrix)
#p = 9.823-11
shapiro.test((resid(lower_anova_c)))
plot(lower_anova_c, which = c(1,2))
hist(resid(lower_anova_c))

##TRANSFORMATION
lower_x <- (lower_matrix$LOWER)^2
x_mod <- aov(lower_x ~ COUNTY, data = lower_matrix)
shapiro.test((resid(x_mod)))
hist(resid(x_mod))

#create ANOVA-month model and assess for normality
lower_anova_m <- aov(LOWER ~ MONTH, data = lower_matrix)
#p = 7.309e-13
shapiro.test((resid(lower_anova_m)))
plot(lower_anova_m, which = c(1,2))
hist(resid(lower_anova_m))
```
### Preliminary Findings
 
Lower quartiles were found to not be normally distributed so a transformation needs to be found or a non-parametric test needs to be considered. 
