---
title: "Identification of Spark Cases"
author: "Maggie Walters"
date: "June 18, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work/")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))

county_vec <- as.character(unique(data$COUNTY))
```

## Document Synopsis

This document aims to identify patterns/trends in the spark cases within each county. The protocol will be as follows:

* Create a matrix (**insert name**) including the first cases within each county as well as the cases the proceed in the following month. 
    + The matrix should include: age at onset, week, district, county, RMSP (just in case)
* Identify those which are in the same district.
* Look for any similarities in age patterns at the beginning of an outbreak. 
```{r Order by week}
#order by week
data <- data[order(data$WEEK),]

#want to find the earliest appearance of a county 
index_mat <- matrix(rep(NA, length(county_vec) * 100), ncol = length(county_vec))
colnames(index_mat) <- county_vec
for(i in 1:length(county_vec)){
  which.county <- which(data$COUNTY == county_vec[i])
  which.county <- which.county[1:100]
  index_mat[,i] <- which.county
}

#find which indeces correspond to which weeks
which_week_mat <- matrix(rep(NA, 3 * max(data$WEEK)), ncol = 3)
colnames(which_week_mat) <- c("WEEK", "MIN", "MAX")
for(i in 1:max(data$WEEK)){
  which.week <- which(data$WEEK == i)
  which_week_mat[i,1] <- i
  which_week_mat[i,2] <- min(which.week, na.rm = TRUE)
  which_week_mat[i,3] <- max(which.week, na.rm = TRUE)
}

#new mat
index_week_mat <- matrix(rep(NA, 3 * length(county_vec) * 100), ncol = 3)
colnames(index_week_mat) <- c("NUM", "COUNTY", "WEEK")

#fill in county
index_week_mat[seq(1:100),2] <- rep(county_vec[1], 100)
for(i in 2: length(county_vec)){
  x <- 100 * (i-1)
  index_week_mat[x + seq(1:100), 2] <- rep(county_vec[i], 100)
}

#fill index
for(i in 1: length(county_vec) * 100){
  index_week_mat[,1] <- index_mat[,]
}

#remove NAs
index_week_mat <- as.data.frame(index_week_mat)
index_week_mat <- subset(index_week_mat, !is.na(index_week_mat$NUM))
which_week_mat <- as.data.frame(which_week_mat)
which_week_mat <- subset(which_week_mat, !is.na(which_week_mat$MIN) & !is.na(which_week_mat$MAX))

#fill week
index_week_mat$NUM <- as.numeric(index_week_mat$NUM)
which_week_mat$MIN <- as.numeric(which_week_mat$MIN)
which_week_mat$MAX <- as.numeric(which_week_mat$MAX)

for(i in 1:length(county_vec) * 100){
  for(j in 1:53){
    x <- index_week_mat$NUM[i] > which_week_mat$MIN[j]
    y <- index_week_mat$NUM[i] < which_week_mat$MAX[j]
    if(x & y){
    index_week_mat[i,3] <- j
     }
  }
}


index_week_mat[1,1]
which(which_week_mat[,2] < index_week_mat[1,1])
which(which_week_mat[,3] > index_week_mat[1,1])

for(i in 1:length(county_vec) * 100){
  if(which(which_week_mat[,2] < index_week_mat[i,1])[1] == which(which_week_mat[,3] > index_week_mat[i,1])[1]){
    index_week_mat[i,3] <- which(which_week_mat[,2] < index_week_mat[i,1])[1]
  }
}

for(i in 1:length(county_vec) * 100){
  which.min.test <- which(which_week_mat[,2] < index_week_mat[i,1])
  which.max.test <- which(which_week_mat[,3] > index_week_mat[i,1])
  if(which.min.test[1] == which.max.test[1]){
    index_week_mat[i,3] <- which.min.test[1]
  }
}

which.min.test <- which(which_week_mat[,2] < index_week_mat[1,1])
  which.max.test <- which(which_week_mat[,3] > index_week_mat[1,1])
  as.numeric(which.min.test)
  as.numeric(which.max.test)
  if(max(which.min.test) == min(which.max.test)){
    index_week_mat[1,3] <- max(which.min.test)
  }


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
