---
title: "Identification of Spark Cases"
author: "Maggie Walters"
date: "June 18, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))
install.packages("devtools")
devtools::install_github("ropenscilabs/arresteddev")
library("arresteddev")

#project set up
setwd("~/Desktop/UGA_REU/lab_work/")
data = read.csv('spm.data.long.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')
data_age = subset(data, !(is.na(data$AGE)))
data_age_confirmed = subset(data_age, (data_age$CONCLUSION=="CONFIRMED" | data_age$CONCLUSION=="CLINICAL"))
data = subset(data_age_confirmed, !(is.na(data_age_confirmed$ONSET)))
data_onset <- as.Date(data$ONSET, "%m/%d/%Y")
mnth <- as.integer(substr(data_onset, 6, 7))
yr<- as.integer(substr(data_onset, 1, 4))

county_vec <- as.character(unique(data$COUNTY))
```

## Document Synopsis

This document aims to identify patterns/trends in the spark cases within each county. The protocol will be as follows:

* Create a matrix (**index_week_mat**) including the first 100 cases within each county as well as the age of that specific case, and the week and county in which it occurred. 
* Identify those which are in the same district.
* Look for any similarities in age patterns at the beginning of an outbreak. 

### Creating index_week_mat

index_week_mat will be created by pulling out the index of the first 100 cases in county *i*. The county name and week in which the case occured will also be recorded. 

```{r First 100 cases in each county, with week and index, echo = FALSE}
index_week_mat <- matrix(rep(NA, 3 * length(county_vec) * 100), ncol = 3)
colnames(index_week_mat) <- c("NUM", "COUNTY", "WEEK")
index_week_mat <- as.data.frame(index_week_mat)

#order by week
data_ordered <- data[order(data$WEEK),]

#matrix with indeces
index_mat <- matrix(rep(NA, length(county_vec) * 100), ncol = length(county_vec))
colnames(index_mat) <- county_vec
for(i in 1:length(county_vec)){
  which.county <- which(data_ordered$COUNTY == county_vec[i])
  which.county <- which.county[1:100]
  index_mat[,i] <- which.county
}

#fill in county
index_week_mat[seq(1:100),2] <- rep(county_vec[1], 100)
for(i in 2: length(county_vec)){
  x <- 100 * (i-1)
  index_week_mat[x + seq(1:100), 2] <- rep(county_vec[i], 100)
}

#fill index
index_week_mat[seq(1:100),1] <- index_mat[,1]
for(i in 2: 387){
  x <- 100 * (i-1)
  index_week_mat[x + seq(1:100),1] <- index_mat[,i]
}

#fill week
index_week_mat <- as.data.frame(index_week_mat)
index_week_mat$WEEK <- as.numeric(index_week_mat$WEEK)
index_week_mat <- subset(index_week_mat, !is.na(index_week_mat$NUM))

for(i in 1:38700){
  x <- index_week_mat$NUM[i]
  x <- as.numeric(as.character(x))
  y <- data_ordered$WEEK[x]
  y <- as.numeric(as.character(y))
  index_week_mat[i,3] <- y
}


```

### Preliminary Analysis:

Some cases do not have more than 100 cases, so they have relatively small case loads. Could change to doing counties with more than 20 cases, which would amount to approximately 80 counties. Again, I am coming across the question of **what is the minimum number of cases necessary for a disease outbreak to be considered an epidemic?**


## Examine ages of first 100 cases and timing

AGE was added as a variable to the index_week_mat in order to examine age distribution patterns. 

```{r add age to index_week_mat}
#fill in age
AGE <- rep(NA, 38700)
index_week_mat <- cbind(AGE, index_week_mat)


for(i in 1:38700){
  x <- index_week_mat[i,2]
  x <- as.numeric(as.character(x))
  index_week_mat[i,1] <- data_ordered$AGE[x]
}

```

## Ordering the cases in ascending order of week of onset

```{r Order by week}
index_week_mat <- index_week_mat[order(index_week_mat$WEEK),] 
```

## Plot age by week for each county

```{r plot age by week}
#need to remove rows 6759 to 38700, just going to make a new matrix called the same thing
which.na <- which(is.na(index_week_mat$WEEK))

index_week_mat <- subset(index_week_mat, !is.na(index_week_mat$NUM) & !is.na(index_week_mat$AGE)
                         & !is.na(index_week_mat$COUNTY) & !is.na(index_week_mat$WEEK))
index_week_mat$WEEK <- as.numeric(index_week_mat$WEEK)
index_week_mat$AGE <- as.numeric(index_week_mat$AGE)
for(i in 1:length(county_vec)){
  x <- subset(index_week_mat, index_week_mat$COUNTY == county_vec[i])
  x$AGE <- as.numeric(as.character(x$AGE))
  plot(x$WEEK, x$AGE, type = "b", main = county_vec[i], xlim=c(0,53))
}
```

### Preliminary analysis:

Graphs of counties with less than 10 cases are really unhelpful visually. Additionally, the the line graph seems to indicate causation between cases, which is not accurate. 

Next step:

* If available, make all of the cases which occurred in the same district the same color. 

## Examination by district

**Just realized that I only have district data for those in Sao Paulo county. I will make graphs for all of the information that I have available.**

```{r examination by district}
x <- subset(data, data$COUNTY == "SAO PAULO")
x <- subset(x, !is.na(x$DISTRICT))
test_district <- as.character(unique(x$DISTRICT))


sp.index_week_mat <- subset(index_week_mat, index_week_mat$COUNTY == county_vec[1])
sp.data <- data[as.numeric(as.character(sp.index_week_mat$NUM)),]
sp.data <- subset(sp.data, !is.na(sp.data$DISTRICT))
unique.district <- unique(sp.data$DISTRICT)

for(i in 1:length(unique.district)){
  dis.which <- which(sp.data$DISTRICT == unique.district[i])
  plot(sp.index_week_mat$WEEK[dis.which], sp.index_week_mat$AGE[dis.which], 
     pch = 19, 
     main = unique.district[i])
}



```

###Preliminary analysis:

It does not seem that analysis via examination by district is going to be very informative. District is only available for Sao Paulo county, and on examination of the first one hundred cases by district for Sao Paulo county, there is not obvious trend. 

## Stage of epidemic in each county during each week of the epidemic
```{r stage of epidemic}

status_mat <- matrix(rep(NA, length(county_vec) * 53), ncol = length(county_vec))
colnames(status_mat) <- county_vec
rownames(status_mat) <- seq(1:53)

for(i in 1:length(county_vec)){
  for(j in 1:53){
    x <- subset(data, data$COUNTY == county_vec[i] & data$WEEK == j)
    status_mat[j,i] <- nrow(x)
  }
}
 
```

* Created the function **active_cases(i, j)**, which describes the amount of infectious cases in one county (*i*) at one time (*j* in weeks). Goal is to scale this value by population size in order to represent the chance that you'd come into contact with an infected individual. 
* These values would also eventually be done to describe the chance of contacting someone from another county who is infected. The chance of being infected by someone in a different county would be scaled by distance from your county to the county in question, and possible immigration rates if possible. 
    + Could also use whether the area is urban or rural as a scaling factor. 
    + Urban centers would be scaled lower for those in rural area because rural to urban is more likely than urban to rural. 
* Also created the function **distance(to, from)** which outputs the distance from one county to another. 


## Population Sizes

Requested three files from Annakate (vac.cov.by.county.csv, pop.by.county.csv, pop.age.by.district.csv). I plan to use the pop.by.county.csv in order to estimate the approxiamate chance of encountering someone with measles.

Also need to estimate the transmissability of measles. R~o~ represents the average number of secondary infections caused by one primary, so R~o~ divided by total contacts should represent the chance that any one contact is one of the ones that is infected. 

**But,** is this really what I am aiming to do here? I want to create something that describes the chance that the original infection case arrived from the county of interest. This should be a representation both of "infectedness" of the county as well as the chance that someone from the county encountered someone from the county of interest. 

This can really be broken down into two parts: the "infectedness" of the county and the county's socialability. Socialability here being how often this county would encounter someone from another county.

**Estimating infectedness of a county:** This can crudely be estimated by the amount of individuals infected at one time over the population size. This is then by definition the proportion of individuals in a county which are infected. 
    + This would be interesting to somehow scale by age. Assuming that mainly adults are immigrating out of rural areas and into urban areas, the infectedness should be higher in counties with a higher age of infection.
    
Re: scaling to age--

* `r x = which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j")`
    + `r scaling factor = mean(index_week_mat$AGE[x]) / max(index_week_mat$AGE[x])`

```{r infectedness of a county}
##contracting from someone in the same district
#i = county_vec[i]
#j = week
#option 1, mean / max
active_cases <- function(i, j){
  x.a <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j - 1")
  sf.a <- mean(index_week_mat$AGE[x]) / max(index_week_mat$AGE[x])
  x.b <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j -2")
  sf.b <- mean(index_week_mat$AGE[x]) / max(index_week_mat$AGE[x])
  x.c <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j - 3")
  sf.c <- mean(index_week_mat$AGE[x]) / max(index_week_mat$AGE[x])
  x.d <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j")
  sf.d <- mean(index_week_mat$AGE[x]) / max(index_week_mat$AGE[x])
  day_one <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (6/7) * status_mat[j-3,i]
  day_two <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (5/7) * status_mat[j-3,i] + sf.d * (1/7) * status_mat[j,i]
  day_three <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (4/7) * status_mat[j-3,i] + sf.d * (2/7) * status_mat[j,i]
  day_four <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (3/7) * status_mat[j-3,i] + sf.d * (3/7) * status_mat[j,i]
  day_five <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (2/7) * status_mat[j-3,i] + sf.d * (4/7) * status_mat[j,i]
  day_six <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (1/7) * status_mat[j-3,i] + sf.d * (5/7) * status_mat[j,i]
  day_seven <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.d * (6/7) * status_mat[j,i]
  days <- c(day_one, day_two, day_three, day_four, day_five, day_six, day_seven)
return(days)
}

#option 2
active_cases.mean <- function(i, j){
  x.a <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j - 1")
  sf.a <- mean(index_week_mat$AGE[x]) 
  x.b <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j -2")
  sf.b <- mean(index_week_mat$AGE[x])
  x.c <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j - 3")
  sf.c <- mean(index_week_mat$AGE[x]) 
  x.d <- which(index_week_mat$COUNTY == county_vec[i] & index_week_mat$WEEK == "j")
  sf.d <- mean(index_week_mat$AGE[x])
  day_one <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (6/7) * status_mat[j-3,i]
  day_two <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (5/7) * status_mat[j-3,i] + sf.d * (1/7) * status_mat[j,i]
  day_three <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (4/7) * status_mat[j-3,i] + sf.d * (2/7) * status_mat[j,i]
  day_four <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (3/7) * status_mat[j-3,i] + sf.d * (3/7) * status_mat[j,i]
  day_five <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (2/7) * status_mat[j-3,i] + sf.d * (4/7) * status_mat[j,i]
  day_six <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.c * (1/7) * status_mat[j-3,i] + sf.d * (5/7) * status_mat[j,i]
  day_seven <- sf.a * 1 * status_mat[j-1,i] + sf.b * 1 * status_mat[j-2,i] + sf.d * (6/7) * status_mat[j,i]
  days <- c(day_one, day_two, day_three, day_four, day_five, day_six, day_seven)
return(days)
}

#what would be the best way to introduce age into this? Should probably scale the days in active_cases. 


pop_size = read.csv('pop.by.county.csv', na.strings = c('', 'NA'), header = TRUE, sep = ',')

prop.active_cases <- function(i, j){
  which.pop <- which(pop_size$COUNTY == county_vec[i])
  prop <- active_cases(i, j) / pop_size$population[which.pop]
  return(prop)
}
```

**Estimating the socialability of a county:** I plan to estimate this using the distance between the two counties. If migration is long-term there should be a term here that increases the socialablity of rural counties because they are sending more, significant people into rural areas. If migration is short-term this should be negated. 

Need some way to scale distance. Essentially, there needs to be some way to say that if the distance between two counties is small, then there is a greater chance that the two counties will come in contact with each other. For now I will use: $$1- (x / \sqrt{y})$$ with x = distance between the two counties and y = area of Sao Paulo state. 

```{r estimating socialability of a county}
install.packages("ggmap")
library(ggmap)

distance_mat <- matrix(rep(NA, length(urban_counties) * length(rural_counties)), ncol = length(urban_counties))
colnames(distance_mat) <- urban_counties
rownames(distance_mat) <- rural_counties

#was going to do this but I ran out of google searches 
for(i in 1:length(urban_counties)){
  for(j in 1:length(rural_counties)){
    distance <- mapdist(from = urban_counties[i], to = rural_counties[j])
    distance_mat[j,i] <- distance$km 
  }
}

distance <- function(k, i){
  to <- county_vec[k]
  from <- county_vec[i]
  distance <- mapdist(from = from, to = to)
  return(distance$km)
}

social <- function(k, i){
  social.prop <- 1 - ((distance(k, i)) / sqrt(248222.8))
  return(social.prop)
}

```

Now I have a function (**prop.active_cases**) that describes the infectedness of a county (not associated with age, but could take it further to this) and a function (**social**) that describes it's socialability (not associated with RMSP). Going to create a function that just relates the two by multiplying them together in order to describe which county the infection most likely spread from.

```{r combine socialability and infectiousness}
#i: index of county spread from
#j: week
#k: index of the county spread to 
spread <- function(i, j, k){
  spread <- prop.active_cases(i, j) * social(k, i)
  return(spread)
}

```

Created the function **spread(i, j, k)** which shows the probability that the infection spread from county *i* to county *k* in week *j*. **spread(i, j, k)** returns a vector of seven values which represent the probability on each day of week *j*.

Now that I have this, if infections were being spread to urban areas from rural areas, the spread outputs should be higher when i is a rural county and when j is an urban county. Important to keep in mind the critical community size for the measles outbreak to occur. Refer back to the Nasell paper (2005).

```{r Urban to Rural}
#RMSP = 1
urban <- subset(data, data$RMSP == "1")
urban_counties <- as.character(unique(urban$COUNTY))

#RMSP = 0 
rural <- subset(data, data$RMSP == "0")
rural_counties <- as.character(unique(rural$COUNTY))


test.mat <- matrix(rep(NA, length(rural_counties) * length(urban_counties)), ncol = length(urban_counties))
#colnames(test.mat) <- c("Day one", "Day two", "Day three",
                        #"Day four", "Day five", "Day six", "Day seven")
colnames(test.mat) <- urban_counties
rownames(test.mat) <- rural_counties

for(i in 1:length(urban_counties)){
  if(i == 34) next
  test.mat[i,] <- spread(which(county_vec == urban_counties[i]), 26, 18)
}

#filling out some of them, but taking a long time to go through. May be of worth to just run the google maps function and find the distance between all of the places and then source this into the function. 
for(i in 1:length(urban_counties)){
  j = 26
  for(k in 1:length(rural_counties)){
    test.mat[k,i] <- spread(which(county_vec == urban_counties[i]), j, which(county_vec == rural_counties[k]))[1]
  }
}

```

### Preliminary Analysis 

I exceeded the Google query limit using the ggmaps tool, so I have to stop on the distance work for now. Going to write out some hypotheses instead. 

* If unvaccinated rural adults were acting as spreaders of the disease to urban areas, then the values when *i* is rural and when *k* is urban should be higher. How would this be quantifiably assessed? I think that a one-tailed t-test may be appropriate, but I'm unsure of whether I'm just fitting what I know to my data. The hypotheses from the t-test would be as follows:

    + H~o~: $\mu ~rural to urban~$ = $\mu ~urban to rural~$
    + H~A~: $\mu ~rural to urban~$ > $\mu ~urban to rural~$

* While this isn't working, I'm going to go ahead and try to make prop.active_cases sensitive to age. 






